Variation in distance between consecutive animal locations

### 2021-03-29 distributions_paper/rudolph_distributions.Rmd
Add text

# Introduction
   Movement is a fundamental ecological process, critical to an animal's biology and it's interactions with the environment. The study of animal movement tends to focus on four basic mechanistic questions: the why, when, where, and how an animal moves, and these questions can be answered at the scales of individuals or populations. A general framework for movement ecology that allows to link movement patterns and processes, starts with a focus on movement itself and the individual's internal state, navigation capacity, motion capacity, and responses to external factors [@nathan2008movement]. The overall goal of studying movement seeks to find the linkages between the causes and mechanisms of animal movement, that result in spatiotemporal patterns that feed back into various ecological and evolutionary processes. Some of these approaches involve separating different movement patterns into behavioral states, with each state having an associated set of parameters that determine that specific movement pattern[@langrock2012flexible].

  Although there is variation in animal movement at the individual scale, the general focus has been on understanding the typical movement patterns, built around average measures at population levels, with the simplest approach assuming model parameters are the same across individuals, known as 'complete pooling' [@langrock2012flexible]. This focus on generality has driven to a lack of frameworks for understanding variation, how it is maintained, and the consequences of this variation across populations, communities, and ecosystems. Some alternatives have been considered to incorporate individual variation with 'no pooling' or 'partial pooling' approaches, where each or some individuals have their own set of parameters [@patterson2009classifying, @jonsen2006robust], however a high number of parameters quickly limits applications to high levels of individual variation. Recent calls for incorporating individual variation in movement [@snell2019consequences, @shaw2020causes] have pushed for the need to develop theory that explores different types of variation in movement patterns to understand its consequences across ecological organizational levels. In particular, when considering species interactions, incorporating individual variation becomes essential to understand how positive or negative feedback loops play into the maintenance of variation in movement as individuals respond to external factors in their environment [@shaw2020causes].

@@

We focused our study of individual variation on the many-banded araçari (*Pterglossus pluricinctus*), a small toucan in Yasuní National Park, Ecuador. Previous studies used radiotelemetry to estimate home ranges and evaluate the potential seed dispersal distances for the *Virola flexuosa* tree [@holbrook2011home, holbrook2007usin]. Over a period of four years, from 2001 to 2005, *P. pluricinctus* individuals were captured and radio-transmitters were attached at the base of the central trail feathers [@holbrook2011home]. Toucan locations were estimated by triangulation using receivers and hand-held antennas, with tracking periods lasting between 4-5 hours per day, locating individual birds every 15 minutes. It is important to note that these methods did not tack all birds simultaneously, and that although individual toucan locations were attempted every 15 minutes, this was not always possible, thus some successive locations have time categories associated to multiples of 15. For our analyses, we focused on toucans with a minimum of **20?** recorded locations, and for each subsequent set of locations we calculated turning angles (in radians), and the associated Euclidean step lengths in meters. We worked under a velocity-based framework, and thus divided step lengths by the time interval during which they were recorded. This gave us a total of **600ish?** data points for angles and velocities.

Describe the data collection from the animov outline. 
Add a section that describes the limitations for this telemetry data. 
From book chapter: Attached radio-transmitters at the base of the central tail feathers. Toucan locations measured by triangulation using receivers and hand-held two-element Yagi antennas. 
Tracking periods laster 4-5 hours with individual birds located every 15 minutes. The 15 minute interval to ensure locations within the time frame of seed passage or regurgitation, which are expected to be longer than 15 minutes.

@@

"Distances travelled per movement bout ranged from 0 to >2000m. Strong leptokurtic distributions of movements with most being <300m. Longest recorded movement for Pluricinctus is 3665m (recorded in 30 minutes)

@@

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

>>>>

# Introduction
   Movement is a fundamental ecological process, critical to an animal's biology and it's interactions with the environment. The study of animal movement tends to focus on four basic mechanistic questions: the why, when, where, and how an animal moves, and these questions can be answered at the scale of individuals or populations. A general framework for movement ecology is needed to link movement patterns and processes. Such a framework would start with a focus on movement itself and the individual's internal state, navigation capacity, motion capacity, and responses to external factors [@nathan2008movement]. The overall goal of studying movement is to find the linkages between the causes and mechanisms of animal movement, that explain spatiotemporal patterns that then feed back into various ecological and evolutionary processes. Some of these approaches involve separating different movement patterns into behavioral states, with each state having an associated set of parameters that determine that specific movement pattern[@langrock2012flexible]. Other approaches have focused on segmenting movement patterns, again in an attempt to associate specific types of movement to underlying behaviors or fine-scale environmental influences (Getz citations, Morales, more focus on the random walk models, Patterson2008). *Indeed, animal movement has consequences.  Such consequences might influence species interactions, ecological function, and vital demographic processes (survival, reproduction, dispersal) that can result in population impacts and evolutionary change (?).*

  Although there is variation in animal movement at the individual scale, the general focus has been on understanding the typical movement patterns, built around average measures at population levels, with the simplest approach assuming model parameters are the same across individuals, known as 'complete pooling' [@langrock2012flexible]. This focus on generality has driven to *a lack of frameworks* for understanding variation, how it is maintained, and the consequences of this variation across populations, communities, and ecosystems. Some alternatives have been considered to incorporate individual variation with 'no pooling' or 'partial pooling' approaches, where each or some individuals have their own set of parameters [@patterson2009classifying, @jonsen2006robust], however a high number of parameters quickly limits applications to high levels of individual variation. Recent calls for incorporating individual variation in movement [@snell2019consequences, @shaw2020causes] have pushed for the need to develop theory that explores different types of variation in movement patterns to understand its consequences across ecological organizational levels. In particular, when considering species interactions, incorporating individual variation becomes essential to understand how positive or negative feedback loops play into the maintenance of variation in movement as individuals respond to external factors in their environment [@shaw2020causes].

@@

We focused our study of individual variation on the many-banded araçari (*Pterglossus pluricinctus*), a small toucan in Yasuní National Park, Ecuador. We used previously collected data from studies using radiotelemetry to estimate home ranges and evaluate the potential seed dispersal distances for the *Virola flexuosa* tree [@holbrook2011home, holbrook2007usin]. Over a period of four years, from 2001 to 2005, *P. pluricinctus* individuals were captured and radio-transmitters were attached at the base of the central tail feathers [@holbrook2011home]. Toucan locations were estimated by triangulation using receivers and hand-held antennas, with tracking periods lasting between 4-5 hours per day, locating individual birds every 15 minutes. Further details on field methods can be found in Holbrook [-@holbrook_home_2011]. It is important to note that these methods did not tack all birds simultaneously, and that although individual toucan locations were attempted every 15 minutes, this was not always possible, thus some successive locations have time categories associated to multiples of 15. For our analyses, we focused on toucans with a minimum of **20?** recorded locations, and we calculated turning angles (in radians), and the associated Euclidean step lengths in meters for each successive set of locations and turning directions. We worked under a velocity-based framework, and thus divided step lengths by the time interval during which they were recorded. This gave us a total of **600ish?** data points for turning angles and velocities.

@@

"Distances travelled per movement bout ranged from 0 to >2000m. Strong leptokurtic distributions of movements with most being <300m. Longest recorded movement for Pluricinctus is 3665m (recorded in 30 minutes)" *this is a corresponding velocity of `r 3665/30` meters per minute.

@@

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(message = FALSE, warning = FALSE, 
                      eval=FALSE,
                      echo = FALSE)
